% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ \textbf{Sets \& Functions}
$\;\to\;$ Proof Techniques
$\;\to\;$ Axiom Systems
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$
$\;\to\;$ Real Analysis
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Logic gave us a precise language. We now need mathematical objects to
reason \emph{about}. Sets are the universe of mathematical objects;
functions are the structure-preserving maps between them.
Together they provide the language of all subsequent mathematics.

\medskip
\noindent\textbf{What this chapter builds.}
We develop naïve and axiomatic set theory: membership, subsets,
operations, power sets, and families. Relations --- including equivalence
relations and partial orders --- give us structure on sets.
Functions capture how sets relate to each other: injections, surjections,
bijections, and cardinality.

\medskip
\noindent\textbf{Where this leads.}
Every algebraic structure is a set with additional operations.
Every proof about $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{R}$ uses
the function and relation vocabulary established here.
Equivalence relations appear again in the integer and rational constructions;
order relations appear throughout analysis.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Structural Roadmap
% =========================================================
\section*{Structural Roadmap}

Each major topic is organised as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Structural Insight}
\end{center}

The global progression is:
\begin{enumerate}
  \item Sets: membership, subsets, operations (union, intersection, difference),
        power sets, the empty set
  \item Relations: binary relations, properties (reflexive, symmetric, transitive),
        composition
  \item Families: indexed families, arbitrary unions and intersections
  \item Equivalence relations: equivalence classes, partitions, quotient sets
  \item Functions: domain, codomain, image, injection, surjection, bijection,
        composition, inverse
  \item Order: partial orders, total orders, well-orders, Zorn's lemma
\end{enumerate}

\vspace{1em}

\begin{remark}[Primary source]
The primary driver is Suppes, \textit{Introduction to Logic},
supplemented by Gerstein, \textit{Introduction to Mathematical Structures
and Proofs}.
\end{remark}

% =========================================================
% Content
% =========================================================
\input{volume-i/sets-relations-functions/notes/index}
