\section{Syntax}

\begin{toolkitbox}{Toolkit}
This section uses truth values, propositional variables, connectives,
well-formed formulas, formation trees, and subformulas to build syntax.
\end{toolkitbox}

\begin{definition}[Model-Theoretic Structure for Propositional Logic]
Let $\mathcal{L}$ be a propositional language over a set of propositional variables $\mathcal{V}$. A \textbf{model} (or \textit{semantic structure}) for $\mathcal{L}$ is defined as a tuple $\mathcal{M} = \langle \mathcal{D}, \mathcal{I} \rangle$, where:
\begin{itemize}
    \item $\mathcal{D}$ is the \textbf{domain of discourse} (or semantic domain). In classical propositional logic, this is the two-element Boolean algebra of truth values, typically denoted as $\mathcal{D} = \{ \top, \bot \}$ (or $\{1, 0\}$).
    \item $\mathcal{I}: \mathcal{V} \to \mathcal{D}$ is the \textbf{interpretation function}, which maps each atomic propositional variable to a truth value in the domain.
\end{itemize}

By the principle of structural induction, the interpretation $\mathcal{I}$ extends uniquely to a total valuation function $\overline{\mathcal{I}}: \text{WFF} \to \mathcal{D}$ over the set of all well-formed formulas. This extension acts as a homomorphism from the algebra of formulas to the Boolean algebra $\mathcal{D}$, satisfying constraints such as:
\[ \overline{\mathcal{I}}(\phi \land \psi) = \overline{\mathcal{I}}(\phi) \sqcap \overline{\mathcal{I}}(\psi) \]
where $\sqcap$ is the meet operation (logical AND) in the Boolean algebra $\mathcal{D}$.

We say the model $\mathcal{M}$ \textbf{satisfies} a formula $\phi$, denoted $\mathcal{M} \models \phi$, if and only if $\overline{\mathcal{I}}(\phi) = \top$.
\end{definition}











\begin{remark*}[Exposition]
The syntax section builds the formal language before assigning it meaning.
It isolates the recursive formation rules for formulas and the structural
notions, such as subformulas and depth, that later support semantic induction
and proof arguments.
\end{remark*}
\begin{definition}[Syntax of Propositional Logic]
The \textbf{syntax} of propositional logic dictates the formal rules for constructing valid expressions.

Syntax is concerned strictly with the structural formation of these strings, completely independent of their semantic meaning or truth values.
\end{definition}
\input{volume-i/propositional-logic/notes/syntax/notes-syntax}
