% BEGIN GENERATED ARTIFACT: prf:normal-form-semantic-classification-propositional-logic
\newpage
\phantomsection
\label{prf:normal-form-semantic-classification-propositional-logic}
\LRAProofFor{thm:normal-form-semantic-classification-propositional-logic}

\begin{remark*}[Return]
\hyperref[thm:normal-form-semantic-classification-propositional-logic]{Return to Theorem (Semantic Classification by Normal Forms).}
\end{remark*}

\begin{theorem*}[Semantic Classification by Normal Forms]
Normal forms give finite certificates for semantic classification. A formula is
satisfiable exactly when some DNF equivalent has a satisfiable term; it is a
contradiction exactly when some DNF equivalent has no satisfiable term; it is a
tautology exactly when some CNF equivalent has no falsifiable clause; and it is
a contingency exactly when it is satisfiable and not a tautology.
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~
TODO: supply the compact proof.
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~
TODO: supply the detailed learning proof.
\end{proof}

\begin{remark*}[Proof structure]
Use logical equivalence to replace \(\varphi\) by an equivalent normal form.
For DNF, apply the DNF satisfiability criterion to detect whether at least one
term can be made true; absence of such a term gives contradiction. For CNF,
use the dual clause check: if no clause can be falsified, every valuation makes
each clause true, so the conjunction is a tautology. Finally, contingency is
the semantic case of satisfiable but not tautological.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:disjunctive-normal-form-propositional-logic]{Disjunctive Normal Form}
  \item \hyperref[def:conjunctive-normal-form-propositional-logic]{Conjunctive Normal Form}
  \item \hyperref[def:satisfiable-formula-propositional-logic]{Satisfiable Formula}
  \item \hyperref[def:contradiction-propositional-logic]{Contradiction}
  \item \hyperref[def:tautology-propositional-logic]{Tautology}
  \item \hyperref[def:contingency-propositional-logic]{Contingency}
  \item \hyperref[thm:dnf-satisfiability-criterion-propositional-logic]{DNF Satisfiability Criterion}
  \item \hyperref[thm:canonical-disjunctive-normal-form-propositional-logic]{Canonical Disjunctive Normal Form Theorem}
  \item \hyperref[thm:canonical-conjunctive-normal-form-propositional-logic]{Canonical Conjunctive Normal Form Theorem}
\end{itemize}
\end{dependencies}

\clearpage
% END GENERATED ARTIFACT: prf:normal-form-semantic-classification-propositional-logic
