% BEGIN GENERATED ARTIFACT: prf:canonical-conjunctive-normal-form-propositional-logic
\newpage
\phantomsection
\label{prf:canonical-conjunctive-normal-form-propositional-logic}
\LRAProofFor{thm:canonical-conjunctive-normal-form-propositional-logic}

\begin{remark*}[Return]
\hyperref[thm:canonical-conjunctive-normal-form-propositional-logic]{Return to Theorem (Canonical Conjunctive Normal Form).}
\end{remark*}

\begin{theorem*}[Canonical Conjunctive Normal Form Theorem]
Every formula \(\varphi\in\WFF\) is logically equivalent to its canonical
conjunctive normal form relative to a fixed ordering of its finitely many
variables.
\[
\varphi\equiv \operatorname{CCNF}_{P_1,\ldots,P_n}(\varphi).
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~
TODO: supply the compact proof.
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~
TODO: supply the detailed learning proof.
\end{proof}

\begin{remark*}[Proof structure]
Use the finite truth-table theorem to list all rows over the variables of
\(\varphi\). For each false row, the corresponding row clause is false exactly
on that row and true on every other row. The conjunction of all such row
clauses is therefore false exactly on the same rows as \(\varphi\). The
coincidence lemma ensures no variables outside the fixed list affect the
comparison.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:canonical-conjunctive-normal-form-propositional-logic]{Canonical Conjunctive Normal Form}
  \item \hyperref[thm:finite-truth-table-propositional-logic]{Finite Truth-Table Theorem}
  \item \hyperref[lem:coincidence-lemma-propositional-logic]{Coincidence Lemma for Propositional Logic}
  \item \hyperref[def:logical-equivalence-propositional-logic]{Logical Equivalence}
\end{itemize}
\end{dependencies}

\clearpage
% END GENERATED ARTIFACT: prf:canonical-conjunctive-normal-form-propositional-logic
