% BEGIN GENERATED ARTIFACT: prf:canonical-disjunctive-normal-form-propositional-logic
\newpage
\phantomsection
\label{prf:canonical-disjunctive-normal-form-propositional-logic}
\LRAProofFor{thm:canonical-disjunctive-normal-form-propositional-logic}

\begin{remark*}[Return]
\hyperref[thm:canonical-disjunctive-normal-form-propositional-logic]{Return to Theorem (Canonical Disjunctive Normal Form).}
\end{remark*}

\begin{theorem*}[Canonical Disjunctive Normal Form Theorem]
Every formula \(\varphi\in\WFF\) is logically equivalent to its canonical
disjunctive normal form relative to a fixed ordering of its finitely many
variables.
\[
\varphi\equiv \operatorname{CDNF}_{P_1,\ldots,P_n}(\varphi).
\]
\end{theorem*}

\begin{proof}
\textbf{Professional Standard Proof.}~
TODO: supply the compact proof.
\end{proof}

\begin{proof}
\textbf{Detailed Learning Proof.}~
TODO: supply the detailed learning proof.
\end{proof}

\begin{remark*}[Proof structure]
Use the finite truth-table theorem to list all rows over the variables of
\(\varphi\). For each true row, the corresponding row conjunction is true
exactly on that row. The disjunction of all such row conjunctions is therefore
true exactly on the same rows as \(\varphi\). The coincidence lemma ensures no
variables outside the fixed list affect the comparison.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:canonical-disjunctive-normal-form-propositional-logic]{Canonical Disjunctive Normal Form}
  \item \hyperref[thm:finite-truth-table-propositional-logic]{Finite Truth-Table Theorem}
  \item \hyperref[lem:coincidence-lemma-propositional-logic]{Coincidence Lemma for Propositional Logic}
  \item \hyperref[def:logical-equivalence-propositional-logic]{Logical Equivalence}
\end{itemize}
\end{dependencies}

\clearpage
% END GENERATED ARTIFACT: prf:canonical-disjunctive-normal-form-propositional-logic
