% volume-i/model-theory/index.tex
% Chapter entry: breadcrumb + roadmap + section inputs

\chapter{Model Theory}

% ── Breadcrumb ────────────────────────────────────────────────

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Proof Techniques
$\;\to\;$ \textbf{Model Theory}
$\;\to\;$ Type Theory
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
The preceding chapters established the syntax and proof theory of
first-order logic: well-formed formulas, deduction rules, and proof
techniques. Predicate logic gave a language; proof techniques gave a
calculus. What has not yet been addressed is \emph{meaning} — the
assignment of mathematical content to logical symbols.

\medskip
\noindent\textbf{What this chapter builds.}
Model theory is the study of the relationship between formal languages
and the mathematical structures that interpret them. The central objects
are $\mathcal{L}$-structures: a domain equipped with interpretations of
every function symbol, relation symbol, and constant in a first-order
language $\mathcal{L}$. From structures, the chapter develops the
satisfaction relation $\mathcal{M} \models \varphi[s]$, which gives
formal meaning to truth-in-a-structure. Elementary equivalence,
substructures, and the compactness theorem are developed in later sections.

\medskip
\noindent\textbf{Where this leads.}
Model theory supplies the semantic foundations for abstract algebra
(a group is an $\mathcal{L}_{\mathrm{gp}}$-structure satisfying the
group axioms), for real analysis (the ordered field axioms characterise
$\mathbb{R}$ up to isomorphism among Archimedean fields), and
ultimately for the ultraproduct constructions that appear in non-standard
analysis and the compactness arguments used throughout the curriculum.
\end{tcolorbox}
\vspace{1em}

% ── Structural Roadmap ────────────────────────────────────────

\subsection*{Structural Roadmap}

Each major topic is organised as:
\begin{center}
\textbf{Definitions $\longrightarrow$ Main Theorems
$\longrightarrow$ Consequences and Structural Insight}
\end{center}

The global progression is:
\begin{enumerate}
  \item \textbf{First-order languages and $\mathcal{L}$-structures.}
        Syntax of languages; domains, function interpretations, relation
        interpretations, constant interpretations.
  \item \textbf{Variable assignments and term evaluation.}
        How terms acquire values in a structure.
  \item \textbf{The satisfaction relation.}
        Inductive definition of $\mathcal{M} \models \varphi[s]$;
        truth and falsity in a structure.
  \item \textbf{Elementary equivalence and substructures.}
        $\mathcal{M} \equiv \mathcal{N}$, $\mathcal{N} \prec \mathcal{M}$,
        the Tarski--Vaught criterion.
  \item \textbf{Theories and models.} \textit{[Planned]}
  \item \textbf{Compactness and L\"{o}wenheim--Skolem.} \textit{[Planned]}
\end{enumerate}

\begin{remark}[Primary sources]
Marker, \textit{Model Theory: An Introduction} (Springer GTM 217),
Chapter~1. Enderton, \textit{A Mathematical Introduction to Logic},
Chapter~2. Hodges, \textit{A Shorter Model Theory}, Chapter~1.
\end{remark}

% ── Section inputs ────────────────────────────────────────────

\section{Languages and Structures}
\input{volume-i/model-theory/notes/index}
