% =========================================================
% Choosing P(n): The Hard Part
% =========================================================
\paragraph{Choosing $P(n)$: The Hard Part.}
\label{sec:choosing-pn}

In textbook problems, the statement $P(n)$ is usually handed to you:
the claim is explicitly ``prove that $\sum_{k=1}^n k = n(n+1)/2$'',
and $P(n)$ is that equality. In original proofs, you must construct
$P(n)$ from the statement of the goal. And sometimes even when $P(n)$
is given, the straightforward choice fails and must be strengthened.

\begin{remark}[Rule 1: $P(n)$ mirrors the logical form of the goal]
Every claim to be proved by induction has the form $\forall n,\; \Phi(n)$
for some predicate $\Phi$. Strip the universal quantifier:
what remains is $P(n)$.

\medskip
\noindent
\begin{tabular}{ll}
\toprule
\textbf{Goal form} & \textbf{Shape of $P(n)$} \\
\midrule
$\forall n,\; f(n) = g(n)$ & Equality: $P(n) :\equiv f(n) = g(n)$ \\[4pt]
$\forall n,\; A(n) \Rightarrow B(n)$ & Conditional: $P(n) :\equiv A(n) \Rightarrow B(n)$ \\[4pt]
$\forall n,\; A(n) \Leftrightarrow B(n)$ & Biconditional: $P(n) :\equiv A(n) \Leftrightarrow B(n)$ \\
\bottomrule
\end{tabular}

\medskip
\noindent
Do not simplify. Do not add conditions. Transcribe the logical form
of $\Phi(n)$ exactly into $P(n)$.
\end{remark}

\begin{remark}[Rule 2: induct on the recursive variable]
When the goal involves a recursively defined operation, identify which
variable the definition recurses on. Induct on that variable and hold
all others fixed as arbitrary.

To find the recursive variable: locate the definition of the operation
and identify which variable is reduced in the recursive clause.

\medskip
\noindent
\begin{tabular}{lll}
\toprule
\textbf{Operation} & \textbf{Recursive clause} & \textbf{Induct on} \\
\midrule
Addition (Peano) & $(n\pp) + m := (n + m)\pp$ & $n$ \\[4pt]
Multiplication (Peano) & $(n\pp) \times m := (n \times m) + m$ & $n$ \\[4pt]
Exponentiation & $m^{n\pp} := m^n \times m$ & $n$ \\
\bottomrule
\end{tabular}
\end{remark}

\begin{remark}[The strengthening principle: when $P(n)$ is too weak]
Sometimes the natural $P(n)$ fails not because the statement is false
but because the inductive hypothesis is insufficient to prove the step.
The fix is counterintuitive: \emph{add more to $P(n)$}, making the claim
stronger.

This works because a stronger $P(n)$ gives you more in the inductive
hypothesis, which is what you actually need to prove $P(n+1)$.
The stronger $P(n)$ is harder to prove overall but easier to propagate.

The signal that strengthening is needed: you write the inductive step,
you need to use some fact about the sequence that $P(n)$ alone does not
provide, and you find yourself thinking ``if only I also knew $P(n-1)$.''
That thought is the signal to switch to strong induction (which implicitly
strengthens $P(n)$ to ``$P(k)$ for all $k \leq n$'').
\end{remark}

\begin{example}[Strengthening: Fibonacci bound requires two prior values]
\textbf{Goal.} Show that the Fibonacci sequence satisfies $F_n < 2^n$
for all $n \geq 1$.

\medskip
\noindent\textbf{Naive attempt.}
Let $P(n)$: ``$F_n < 2^n$.''

Inductive step: assume $F_n < 2^n$; show $F_{n+1} < 2^{n+1}$.

We have $F_{n+1} = F_n + F_{n-1}$. We know $F_n < 2^n$ by the
inductive hypothesis. But what do we know about $F_{n-1}$?
Nothing: $P(n)$ says nothing about $F_{n-1}$. The step is stuck.

\medskip
\noindent\textbf{Fix: strong induction gives two terms.}
Use the strong inductive hypothesis: $F_k < 2^k$ for all $k \leq n$.
Then:
\[
F_{n+1} = F_n + F_{n-1} < 2^n + 2^{n-1} < 2^n + 2^n = 2^{n+1}.
\]
The step completes because strong induction provides $F_{n-1} < 2^{n-1}$.
This is the canonical signal that strong induction is needed: the
recursive step requires knowing $P(k)$ for some $k < n$ other than
$k = n-1$.
\end{example}

\begin{remark}[Vacuous base cases in conditional statements]
When $P(n)$ is a conditional and its hypothesis is false at the base
value, the base case is vacuously true. This is expected and requires
no further argument.

For example, if the claim is ``for all $n \geq 1$, if $n > 0$ and
$m > 0$ then $nm > 0$'' and you induct starting at $n = 0$, then
$P(0)$: ``if $0 > 0$ and $m > 0$ then $0 \times m > 0$'' has a
false hypothesis ($0 > 0$ is false), so $P(0)$ is vacuously true.

This signals that the claim only has content for values of $n$ where
the hypothesis holds. The induction then establishes the claim for
all $n$ where the hypothesis is satisfied.
\end{remark}

\begin{remark}[Diagnostic questions when stuck choosing $P(n)$]
\begin{enumerate}[itemsep=3pt]
  \item What exactly do I need to prove for $n = 0, 1, 2, 3$?
    Write it out. The pattern of $P(n)$ will appear.
  \item Does the inductive step require knowing $P(n-1)$ as well as
    $P(n)$? If yes, switch to strong induction.
  \item Does the inductive step produce a bound or equation slightly
    weaker than $P(n+1)$? If yes, strengthen $P(n)$.
  \item After writing $P(n\pp)$, can I apply the recursive definition
    to expose $P(n)$? If not, the wrong variable may have been chosen.
\end{enumerate}
\end{remark}
