% =========================================================
% Statement Forms → Proof Strategies Map
% =========================================================
\paragraph{Statement Forms and Their Proof Strategies.}

The following table maps every common statement form to the proof
strategy it demands. This is not a list of suggestions --- it is a
mechanical lookup. Once you have identified the outermost logical
form of the statement you are trying to prove, the table tells you
how to open your proof.

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{p{0.28\textwidth} p{0.30\textwidth} p{0.32\textwidth}}
\toprule
\textbf{Statement Form} & \textbf{Proof Strategy} & \textbf{Opening Move} \\
\midrule
$\forall x,\; P(x)$ &
  Introduce arbitrary element &
  ``Let $x$ be arbitrary.'' \\

$P \Rightarrow Q$ &
  Direct proof &
  ``Assume $P$.'' \\

$P \Rightarrow Q$ &
  Contrapositive &
  ``Assume $\neg Q$.'' \\

$P \Rightarrow Q$
(when negation more useful) &
  Contradiction &
  ``Assume $P$ and $\neg Q$.'' \\

$P \Leftrightarrow Q$ &
  Two separate directions &
  Prove $P \Rightarrow Q$, then $Q \Rightarrow P$. \\

$\exists x,\; P(x)$ &
  Construct a witness &
  ``Define $x := \ldots$ and verify $P(x)$.'' \\

$\exists!\, x,\; P(x)$ &
  Existence then uniqueness &
  Construct witness; then assume-two-and-compare. \\

$A = B$ (sets) &
  Double inclusion &
  Prove $A \subseteq B$ and $B \subseteq A$. \\

$a = b$ (algebraic elements) &
  Satisfy-and-cite or algebra &
  Show $a$ satisfies the defining property of $b$. \\

$P(n)$ for all $n \in \mathbb{N}$ &
  Induction &
  Base case; inductive step. \\

``$X$ is unique'' &
  Assume-two-and-compare &
  ``Let $x, y$ both satisfy the definition.'' \\
\bottomrule
\end{tabular}
\end{center}

\begin{remark}[Reading the table: defaults and alternatives]
Some statement forms admit more than one strategy, and the table
reflects this with multiple rows for the same form.

For $P \Rightarrow Q$, the default is direct proof: assume $P$ and
derive $Q$ by forward reasoning. Contrapositive is reached for when
$\neg Q$ is more useful than $P$ as an assumption --- this happens
when the negation of the conclusion is a strong, concrete condition
(for example, ``$n$ is odd'' is concrete; ``$n$ is not even'' is the
same thing but sounds weaker). Contradiction is reached for when neither
$P$ alone nor $\neg Q$ alone gets you far, but combining them produces
a quick absurdity.

The hierarchy is: try direct proof first. If the hypothesis is hard
to use forward, try contrapositive. If neither works, try contradiction.
Do not reach for contradiction as a default just because the proof
feels difficult.
\end{remark}

\begin{remark}[Set equality versus element equality are genuinely different]
The two equality rows in the table look similar but demand completely
different techniques, and conflating them is a common error.

$A = B$ as \emph{sets} is proved by double inclusion because sets are
defined extensionally: two sets are equal if and only if they have the
same elements. Proving $A = B$ by any other means --- saying they are
``clearly the same'' or ``follow from the definition'' --- is not a proof.
The double inclusion must be made explicit.

$a = b$ as \emph{elements in an algebraic structure} is different.
Here $b$ is typically a uniquely-determined object: an inverse, an
identity, a limit, a fixed point. The strategy is to show $a$ satisfies
the same defining property as $b$, and then invoke the previously proved
uniqueness theorem to conclude $a = b$. This is the satisfy-and-cite pattern
(Section~\ref{sec:satisfy-cite}).

The mistake is trying to use double inclusion on algebraic elements, or
trying to use algebraic manipulation on sets. The form of the objects
determines which technique applies.
\end{remark}

\begin{remark}[The biconditional is always two separate proofs]
A biconditional $P \Leftrightarrow Q$ is \emph{never} proved in a single
argument that runs from $P$ to $Q$ and ``loops back.'' It is always
proved in two distinct proofs:

\medskip
\noindent \textbf{($\Rightarrow$)} Assume $P$. Prove $Q$.

\noindent \textbf{($\Leftarrow$)} Assume $Q$. Prove $P$.

\medskip
\noindent
Label the two directions explicitly. The two directions often require
completely different techniques and different amounts of work.
In characterisation theorems in analysis, one direction is typically
a two-line unraveling of a definition while the other requires a
nontrivial inequality argument. Treating them as one argument is how
gaps get hidden.
\end{remark}
