% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ \textbf{Predicate Calculus}
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ Axiom Systems
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$
$\;\to\;$ Real Analysis
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Propositional logic gave us a language for combining truth values with
connectives. But it cannot express ``every integer has a successor'' or
``there exists a prime greater than 100'' --- statements that quantify
over objects. Predicate logic provides the missing machinery.

\medskip
\noindent\textbf{What this chapter builds.}
We extend propositional syntax with variables, constants, function
symbols, predicate symbols, and quantifiers.
The semantics are now structures: a domain of objects plus
interpretations of every symbol.
The proof theory adds quantifier introduction and elimination rules.
Soundness and completeness (Gödel's theorem) are the major metatheorems.

\medskip
\noindent\textbf{Where this leads.}
Every subsequent chapter uses predicate logic as its underlying language.
The axiomatic construction of $\mathbb{N}$ is a predicate-logic theory.
Set theory, ring theory, and topology are all first-order theories.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Structural Roadmap
% =========================================================
\section*{Structural Roadmap}

The same four-layer architecture as propositional logic applies:
\begin{center}
\textbf{Syntax $\longrightarrow$ Semantics
$\longrightarrow$ Proof Theory $\longrightarrow$ Metatheory}
\end{center}

The global progression is:
\begin{enumerate}
  \item Syntax: terms, atomic formulas, well-formed formulas,
        free and bound variables, substitution
  \item Semantics: structures, variable assignments, satisfaction,
        validity, logical equivalence, substitution lemmas
  \item Quantifiers: universal, existential, unique existential,
        bounded quantifiers, negation, commutation, prenex normal form
  \item Proof theory: quantifier rules (UI, UG, EI, EG),
        equality, soundness and completeness
  \item Translation: English to predicate logic, scope, square of opposition
  \item Reference: quantifier fallacies, comparison tables
\end{enumerate}

\vspace{1em}

\begin{remark}[Primary source]
The primary driver is Bjørndahl's \textit{Logic and Proof},
supplemented by Gerstein's \textit{Introduction to Mathematical Structures
and Proofs}.
\end{remark}

% =========================================================
% Content
% =========================================================
\input{volume-i/predicate-logic/notes/index}
