% BEGIN GENERATED TOPIC: normal-forms-conversion-procedures
\begin{remark*}[Equivalence-preserving conversion]
Normal form conversion is a rewrite process that replaces a formula by a
logically equivalent formula with a prescribed syntactic shape. Each local
rewrite is justified by an equivalence law, and each replacement inside a
larger formula is justified by
\hyperref[thm:replacement-of-logical-equivalents-propositional-logic]{Replacement of Logical Equivalents}.

The goal is not to preserve the exact parse tree of the original formula. The
goal is to preserve truth value under every valuation while moving the formula
toward
\hyperref[def:negation-normal-form-propositional-logic]{negation normal form},
\hyperref[def:disjunctive-normal-form-propositional-logic]{disjunctive normal form},
or
\hyperref[def:conjunctive-normal-form-propositional-logic]{conjunctive normal form}.
\end{remark*}

\begin{remark*}[Step 1: eliminate biconditionals]
Replace every biconditional by an equivalent formula using the
\hyperref[thm:biconditional-equivalence-laws-propositional-logic]{Biconditional Equivalence Laws}.
A standard replacement is
\[
\alpha\leftrightarrow\beta
\quad\rightsquigarrow\quad
(\alpha\rightarrow\beta)\land(\beta\rightarrow\alpha).
\]
After this step, no occurrence of \(\leftrightarrow\) remains.
\end{remark*}

\begin{remark*}[Step 2: eliminate conditionals]
Replace every conditional by an equivalent disjunction using the
\hyperref[thm:conditional-equivalence-laws-propositional-logic]{Conditional Equivalence Laws}.
The standard replacement is
\[
\alpha\rightarrow\beta
\quad\rightsquigarrow\quad
\neg\alpha\lor\beta.
\]
After this step, the only connectives still present are \(\neg\), \(\land\),
and \(\lor\).
\end{remark*}

\begin{remark*}[Step 3: push negations inward]
Use
\hyperref[thm:de-morgan-laws-propositional-logic]{De Morgan Laws for Propositional Logic}
and double-negation simplification to move every negation down to the literal
level:
\[
\neg(\alpha\land\beta)\rightsquigarrow \neg\alpha\lor\neg\beta,
\qquad
\neg(\alpha\lor\beta)\rightsquigarrow \neg\alpha\land\neg\beta,
\qquad
\neg\neg\alpha\rightsquigarrow \alpha.
\]
When this step terminates, the result is in negation normal form: negations
occur only immediately in front of propositional variables.
\end{remark*}

\begin{remark*}[Step 4A: distribute for DNF]
To obtain DNF from negation normal form, distribute conjunction over
disjunction using the
\hyperref[thm:distributive-laws-propositional-logic]{Distributive Laws for Propositional Logic}:
\[
\alpha\land(\beta\lor\gamma)
\quad\rightsquigarrow\quad
(\alpha\land\beta)\lor(\alpha\land\gamma),
\]
and similarly
\[
(\beta\lor\gamma)\land\alpha
\quad\rightsquigarrow\quad
(\beta\land\alpha)\lor(\gamma\land\alpha).
\]
Repeatedly applying these rewrites moves disjunctions outward until the formula
is a finite disjunction of conjunction terms.
\end{remark*}

\begin{remark*}[Step 4B: distribute for CNF]
To obtain CNF from negation normal form, distribute disjunction over
conjunction:
\[
\alpha\lor(\beta\land\gamma)
\quad\rightsquigarrow\quad
(\alpha\lor\beta)\land(\alpha\lor\gamma),
\]
and similarly
\[
(\beta\land\gamma)\lor\alpha
\quad\rightsquigarrow\quad
(\beta\lor\alpha)\land(\gamma\lor\alpha).
\]
Repeatedly applying these rewrites moves conjunctions outward until the formula
is a finite conjunction of clauses.
\end{remark*}

\begin{remark*}[DNF conversion workflow]
To convert a formula to DNF:
\begin{enumerate}
  \item eliminate biconditionals;
  \item eliminate conditionals;
  \item push negations inward to obtain negation normal form;
  \item distribute conjunction over disjunction until the result is a finite
        disjunction of elementary conjunctions.
\end{enumerate}
The resulting formula is logically equivalent to the original formula, but it
need not be the shortest or only possible DNF representation.
\end{remark*}

\begin{remark*}[CNF conversion workflow]
To convert a formula to CNF:
\begin{enumerate}
  \item eliminate biconditionals;
  \item eliminate conditionals;
  \item push negations inward to obtain negation normal form;
  \item distribute disjunction over conjunction until the result is a finite
        conjunction of clauses.
\end{enumerate}
The resulting formula is logically equivalent to the original formula, but it
need not be the shortest or only possible CNF representation.
\end{remark*}

\begin{remark*}[Termination]
The first three stages terminate because they remove or strictly simplify
specific connective patterns. Eliminating biconditionals reduces the number of
\(\leftrightarrow\) symbols. Eliminating conditionals reduces the number of
\(\rightarrow\) symbols. Pushing negations inward reduces the distance between
each negation and the propositional variables it governs, while double
negations shorten the formula.

The distribution stages terminate when guided by a target measure: for DNF,
keep applying distributions that move \(\lor\) outward past \(\land\); for CNF,
keep applying distributions that move \(\land\) outward past \(\lor\). Each
application lowers the number of target-obstructing connective patterns, even
though it may duplicate subformulas.
\end{remark*}

\begin{remark*}[Correctness]
Correctness is by equivalence preservation. Each rewrite instance is one of the
listed logical equivalence laws:
\hyperref[thm:biconditional-equivalence-laws-propositional-logic]{biconditional laws},
\hyperref[thm:conditional-equivalence-laws-propositional-logic]{conditional laws},
\hyperref[thm:de-morgan-laws-propositional-logic]{De Morgan laws}, or
\hyperref[thm:distributive-laws-propositional-logic]{distributive laws}.
Replacement of logical equivalents permits the rewrite to occur inside an
arbitrary larger formula. Therefore the final normal form is logically
equivalent to the original formula.
\end{remark*}

\begin{remark*}[Non-uniqueness]
Ordinary DNF and CNF are not unique. Reordering terms or clauses, reordering
literals inside a term or clause, duplicating a term, deleting a redundant
term, or applying absorption-style simplifications can produce different
formulas with the same truth table and the same normal-form type.

Canonical normal forms impose extra conventions to remove this freedom. The
ordinary conversion procedures in this topic do not impose those conventions.
\end{remark*}

\begin{remark*}[Size blowup]
Distribution can duplicate subformulas. As a result, converting a compact
formula to DNF or CNF can produce a formula whose size is much larger than the
original. This size blowup is a syntactic cost of forcing a formula into a
flat disjunction-of-terms or conjunction-of-clauses shape.
\end{remark*}

\begin{remark*}[Comparison]
\[
\begin{array}{c|c|c}
\text{Target form} & \text{Final outer connective pattern} & \text{Distribution used}\\
\hline
\text{NNF} & \text{negations only at literals} & \text{none}\\
\text{DNF} & \text{disjunction of conjunction terms} &
\land\text{ over }\lor\\
\text{CNF} & \text{conjunction of clauses} &
\lor\text{ over }\land
\end{array}
\]
All three workflows use equivalence-preserving rewrites. DNF and CNF share the
same first three stages and differ only in the final distribution direction.
\end{remark*}
% END GENERATED TOPIC: normal-forms-conversion-procedures
