% =========================================================
% Proof Techniques — Notes Assembly
% =========================================================




\begin{tcolorbox}[
  colback=gray!6, colframe=gray!40, arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic $\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ \textbf{Proof Techniques}
$\;\to\;$ Real Analysis $\;\to\;$ Algebraic Structures
$\;\to\;$ Linear Algebra $\;\to\;$ $\cdots$
\end{center}
\medskip
\noindent\textbf{How we got here.}
You have spent the first part of this curriculum learning to reason
precisely. Propositional logic gave you the algebra of truth and the laws
of valid inference. Predicate logic extended that to statements about
objects: what it means to say ``for all $x$'' or ``there exists $y$.''
Sets and functions gave you the universe of objects to reason about.
Every tool you will ever use in a mathematical proof is ultimately grounded
in those foundations.

\medskip
\noindent\textbf{What this chapter builds.}
This chapter answers a different question: \emph{how do you actually write
a proof?} Knowing the laws of logic is not the same as knowing how to sit
down in front of a blank page, read a mathematical statement, and construct
a valid argument. This chapter makes that process explicit and systematic.
You will learn to read a statement and immediately identify its proof
strategy; to execute that strategy step by step without losing track of
what you have established; to use the three fundamental proof structures
in the situations where each is appropriate; to deploy induction correctly
and confidently; and to recognize the handful of algebraic tactics that
appear in virtually every proof in abstract mathematics.

\medskip
\noindent\textbf{Where this leads.}
Every proof you write from this point forward uses the machinery developed
here. In real analysis, convergence and continuity are proved using direct
proof and contradiction. In abstract algebra, uniqueness results use the
assume-two-and-compare pattern and equalities use satisfy-and-cite.
Induction threads through number theory, algebra, and combinatorics.
The patterns are not numerous, but they are universal: master them here
and nothing in the later curriculum will feel structurally unfamiliar.
\end{tcolorbox}
\vspace{1em}

\section*{Structural Roadmap}

This chapter is organized in six sections of increasing specificity.

\begin{center}
\textbf{Architecture $\longrightarrow$ Construction
$\longrightarrow$ Structures $\longrightarrow$ Induction
$\longrightarrow$ Tactics $\longrightarrow$ Reference
$\longrightarrow$ Analysis Preview}
\end{center}

\medskip
\noindent
\textbf{Sections 1--4} should be read in order on your first pass.
They form a complete, coherent account of how proofs are constructed
from start to finish. Section~1 answers the question of \emph{what kind
of argument} a given statement calls for. Section~2 shows you how to
execute that argument \emph{line by line}. Section~3 goes deep on the
five structural patterns that cover almost every proof you will write.
Section~4 gives induction the full treatment it deserves: not just the
mechanical template, but the reasoning behind it, the different variants,
and the mistakes that derail students who never understood why it works.

\medskip
\noindent
\textbf{Sections 5--6} are reference material. Read Section~5 (Tactics)
once, carefully, when you encounter an algebraic proof that stalls at a
specific step. After that, treat it as a lookup table. Section~6 is a
pure reference: a map from ``stuck situations'' to strategies.

\medskip
\noindent
\textbf{Section 7} is a forward-looking preparation for Volume~III.
It does not teach real analysis; it equips you with a framework so that
when you arrive at sequences and metric spaces, the proof patterns will
not feel alien. It introduces the six structural types that every analysis
exercise falls into, the twelve canonical proof moves used throughout
metric space theory, and the four-layer hierarchy that organizes those
moves. Read it after completing Sections~1--4 and before beginning Volume~III.

\begin{remark}[Four distinct questions that students conflate]
The most common source of proof-writing paralysis is conflating four
questions that have different answers and require different tools.
\textbf{Architecture} asks: what kind of proof is this, and what skeleton
should it have? \textbf{Construction} asks: once the skeleton is chosen,
how do I execute it line by line? \textbf{Structures} asks: what do the
standard proof patterns look like in practice? \textbf{Tactics} asks:
once the proof is in progress, which specific algebraic moves are
available? A student who asks ``how do I start?'' may be asking any of
the first three questions simultaneously. Separating them is the first
act of a practised proof writer.
\end{remark}

\input{volume-i/proof-techniques/notes/architecture/index}
\input{volume-i/proof-techniques/notes/construction/index}
\input{volume-i/proof-techniques/notes/structures/index}
\input{volume-i/proof-techniques/notes/induction/index}
\input{volume-i/proof-techniques/notes/tactics/index}
\input{volume-i/proof-techniques/notes/reference/index}
\input{volume-i/proof-techniques/notes/analysis-preview/index}
