\subsubsection*{Semantic Justification of Argument Forms}

\begin{remark*}[Working vocabulary]
This topic uses tautology, logical consequence, valid argument forms, and
truth-table reasoning to connect named inference patterns with semantic
validity.
\end{remark*}

% BEGIN GENERATED ARTIFACT: thm:tautological-implication-valid-argument-form-propositional-logic
\begin{tcolorbox}[colback=thmbox, colframe=thmborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Theorem (Tautological Implication Gives a Valid Argument Form)}},
  fonttitle=\small\bfseries]
\begin{theorem}[Tautological Implication Gives a Valid Argument Form]
\label{thm:tautological-implication-valid-argument-form-propositional-logic}
Let \(\Gamma=\{\gamma_1,\ldots,\gamma_n\}\subseteq\WFF\) and let
\(\varphi\in\WFF\). If
\[
(\gamma_1\land\cdots\land\gamma_n)\to\varphi
\]
is a tautology, then the argument form \((\Gamma,\varphi)\) is valid.

\noindent\hyperref[prf:tautological-implication-valid-argument-form-propositional-logic]{Go to proof.}
\end{theorem}
\end{tcolorbox}

\begin{remark*}[Standard quantified statement]
\[
\operatorname{Tautology}((\gamma_1\land\cdots\land\gamma_n)\to\varphi)
\Longrightarrow
\operatorname{ValidArgumentForm}(\{\gamma_1,
\ldots,\gamma_n\},\varphi).
\]
\end{remark*}

\begin{remark*}[Predicate reading]
If the conditional from the conjunction of all premises to the conclusion is
always true, then the premises cannot all be true while the conclusion is false.
\end{remark*}

\begin{remark*}[Interpretation]
This theorem converts a tautology check into an argument-validity check. It is
one reason truth tables can validate named inference rules.
\end{remark*}

\begin{remark*}[Exposition]
The semantic proof pattern for a valid argument form is always the same. Let an
arbitrary truth assignment make all premises true. Then show that the conclusion
must also be true. If this can be done for an arbitrary truth assignment, then no
counterexample assignment exists.
\end{remark*}

\begin{remark*}[Exposition]
This section does not yet define formal derivations. The word valid here is
semantic: it refers to preservation of truth under truth assignments. Later proof
systems will turn selected valid patterns into formal rules of inference.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:tautology-propositional-logic]{Tautology}
  \item \hyperref[def:valid-argument-form-propositional-logic]{Valid Argument Form}
  \item \hyperref[def:logical-consequence-propositional-logic]{Logical Consequence}
\end{itemize}
\end{dependencies}
% END GENERATED ARTIFACT: thm:tautological-implication-valid-argument-form-propositional-logic
