\newpage
\phantomsection
\label{prf:euclid-i-32}
\LRAProofFor{thm:euclid-i-32}

\begin{remark*}[Return]
\hyperref[thm:euclid-i-32]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Euclid I.32: Triangle angle sum]
In any triangle, if one of the sides is produced, then the exterior angle is
equal to the two interior and opposite angles, and the three interior angles of
the triangle are equal to two right angles.
\end{theorem*}

\input{volume-i/book-geometry/euclidean-geometry/notes/compass-and-straightedge-constructions/figure-euclid-i-32}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
Let $ABC$ be a triangle. Through $C$, draw a line parallel to $AB$ by Euclid
I.31. The straight line $AC$ cuts the two parallels, so the angle made by $AC$
with the parallel through $C$ equals the interior angle at $A$. Likewise, the
straight line $BC$ cuts the two parallels, so the angle made by $BC$ with the
parallel through $C$ equals the interior angle at $B$.

At the point $C$, these two copied angles together with the original angle
$ACB$ form a straight angle, that is, two right angles. Substituting the equal
alternate angles for the angles at $A$ and $B$, the three interior angles of
triangle $ABC$ are equal to two right angles. The same equal-angle comparison
also identifies the exterior angle with the sum of the two opposite interior
angles.
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart
The diagram places a parallel to the base through the top vertex. The base
angle at $A$ reappears at $C$ on one side of the triangle, and the base angle
at $B$ reappears at $C$ on the other side. Along the parallel line through
$C$, those two copied angles and the original vertex angle exactly fill a
straight line. A straight line has two right angles, so the three angles of the
triangle sum to two right angles.
\end{proof}

\begin{remark*}[Proof structure]
Draw a parallel through one vertex, transfer the two remote angles by alternate
interior angles, and read the three angles as a straight angle.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:euclid-triangle-angle-classes]{Triangle angle classes}
  \item \hyperref[thm:euclid-i-31]{Euclid I.31}
  \item \hyperref[ax:euclid-postulate-parallel]{Euclid Postulate V}
\end{itemize}
\end{dependencies}

\clearpage
