\newpage
\phantomsection
\label{prf:euclid-i-31}
\LRAProofFor{thm:euclid-i-31}

\begin{remark*}[Return]
\hyperref[thm:euclid-i-31]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Euclid I.31: Drawing a parallel through a point]
Through a given point, to draw a straight line parallel to a given straight
line.
\end{theorem*}

\input{volume-i/book-geometry/euclidean-geometry/notes/compass-and-straightedge-constructions/figure-euclid-i-31}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
Let $P$ be the given point and let $AB$ be the given straight line. Draw a
transversal from $P$ to a point on $AB$. At $P$, construct an angle equal to
the alternate interior angle made by the transversal with $AB$, and draw the
straight line through $P$ forming that equal angle.

The two lines cut by the transversal have equal alternate interior angles.
Euclid's parallel criterion, which rests on the parallel postulate, gives that
the constructed line through $P$ is parallel to $AB$.
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart
The construction copies an angle rather than guessing a direction. Once the
angle at $P$ matches the alternate angle at the given line, the transversal
sees the same slant at both intersections. In Euclidean geometry that equality
of alternate interior angles is exactly the signal that the two straight lines
do not meet when produced.
\end{proof}

\begin{remark*}[Proof structure]
Draw a transversal, copy the alternate angle at the given point, and invoke the
parallel criterion to identify the constructed line as parallel to the given
line.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:euclid-parallel-straight-lines]{Parallel straight lines}
  \item \hyperref[ax:euclid-postulate-draw-straight-line]{Euclid Postulate I}
  \item \hyperref[ax:euclid-postulate-parallel]{Euclid Postulate V}
\end{itemize}
\end{dependencies}

\clearpage
