\newpage
\phantomsection
\label{prf:euclid-i-46}
\LRAProofFor{thm:euclid-i-46}

\begin{remark*}[Return]
\hyperref[thm:euclid-i-46]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Euclid I.46: Constructing a square on a segment]
On a given straight line, to describe a square.
\end{theorem*}

\input{volume-i/book-geometry/euclidean-geometry/notes/compass-and-straightedge-constructions/figure-euclid-i-46}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
Let $AB$ be the given segment. At $A$, draw a perpendicular to $AB$ and cut off
$AD$ equal to $AB$. Through $D$, draw a line parallel to $AB$; through $B$,
draw a line parallel to $AD$. Let those two parallels meet at $C$.

The quadrilateral $ABCD$ has opposite sides parallel, so it is a parallelogram.
The construction gives one right angle, and the parallel lines transfer right
angles to the remaining corners. Also $AD=AB$ by construction, and opposite
sides of the parallelogram are equal. Thus all four sides are equal and all
four angles are right angles. By the definition of a square, $ABCD$ is the
square on $AB$.
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart
The construction first creates one side perpendicular to the given segment and
equal to it. The two parallels then close the quadrilateral without changing
the directions: horizontal sides stay parallel to $AB$, and vertical sides stay
parallel to $AD$. This makes a right-angled parallelogram whose adjacent sides
are equal, hence a square.
\end{proof}

\begin{remark*}[Proof structure]
Construct a perpendicular equal side, close the figure with parallels, and use
parallel transfer of right angles plus equality of opposite sides to identify a
square.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:euclid-quadrilateral-classes]{Square}
  \item \hyperref[thm:euclid-i-31]{Euclid I.31}
  \item \hyperref[ax:euclid-postulate-draw-straight-line]{Euclid Postulate I}
  \item \hyperref[ax:euclid-postulate-right-angles-equal]{Euclid Postulate IV}
\end{itemize}
\end{dependencies}

\clearpage
