\newpage
\phantomsection
\label{prf:euclid-i-47}
\LRAProofFor{thm:euclid-i-47}

\begin{remark*}[Return]
\hyperref[thm:euclid-i-47]{Return to Theorem}
\end{remark*}

\begin{theorem*}[Euclid I.47: Pythagorean theorem]
In right-angled triangles, the square on the side subtending the right angle is
equal to the squares on the sides containing the right angle.
\end{theorem*}

\begin{proof}[Professional Standard Proof]
\LRAProofBodyStart
\input{volume-i/book-geometry/euclidean-geometry/notes/compass-and-straightedge-constructions/figure-euclid-i-47}

Let $ABC$ be a right-angled triangle with right angle at $A$. Construct the
squares on $AB$, $AC$, and $BC$ by Euclid I.46. Through $A$ draw the line
parallel to the appropriate sides of the square on $BC$, dividing that square
into two rectangles.

By Euclid I.4, the triangles built from the side $AB$ and the corresponding
side of the hypotenuse-square are congruent in the required side-angle-side
comparison. By Euclid I.41, the square on $AB$ is double one of those
triangles, and the matching rectangle in the hypotenuse-square is double the
congruent triangle. Hence the square on $AB$ equals that rectangle. The same
argument on the side $AC$ shows that the square on $AC$ equals the other
rectangle in the hypotenuse-square.

The two rectangles together make up the square on $BC$. Therefore, by Common
Notion II, the square on $AB$ together with the square on $AC$ equals the
square on $BC$.
\end{proof}

\begin{proof}[Detailed Learning Proof]
\LRAProofBodyStart
Euclid's proof is an area comparison, but not an algebraic computation of
areas as real numbers. The square on the hypotenuse is split into two
rectangles. One rectangle is shown equal to the square on one leg; the other
rectangle is shown equal to the square on the other leg. Since the two
rectangles together make up the square on the hypotenuse, the square on the
hypotenuse equals the sum of the two leg-squares.

This matters for the later trigonometric construction. When a right triangle
is normalized so that its hypotenuse has unit length, Euclid I.47 is the
synthetic source of the identity
\[
  \cos^2\theta+\sin^2\theta=1.
\]
The coordinate and area-algebra version of this proof belongs later, after
Cartesian coordinates and numerical area have been licensed.
\end{proof}

\begin{remark*}[Dependency status]
This is the Euclidean windmill proof at the level needed for the trigonometric
construction. The local spine now records the congruence, parallel, area, and
square-construction tools that the proof uses.
\end{remark*}

\begin{remark*}[Proof structure]
Construct squares on the three sides of a right triangle, split the square on
the hypotenuse into two rectangles, prove each rectangle equal to the
corresponding square on a leg, and add the two equalities.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:euclid-right-angle-and-perpendicular]{Right angle and perpendicular}
  \item \hyperref[def:euclid-triangle-angle-classes]{Right-angled triangle}
  \item \hyperref[def:euclid-quadrilateral-classes]{Square}
  \item \hyperref[ax:euclid-postulate-draw-straight-line]{Euclid Postulate I}
  \item \hyperref[ax:euclid-common-notion-add-equals]{Euclid Common Notion II}
  \item \hyperref[ax:euclid-common-notion-subtract-equals]{Euclid Common Notion III}
  \item \hyperref[thm:euclid-i-4]{Euclid I.4}
  \item \hyperref[thm:euclid-i-41]{Euclid I.41}
  \item \hyperref[thm:euclid-i-46]{Euclid I.46}
\end{itemize}
\end{dependencies}

\clearpage
