% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
\textbf{Propositional Logic}
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ Axiom Systems
$\;\to\;$ $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$
$\;\to\;$ Real Analysis
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
This is the starting point. Mathematics needs a precise language for
reasoning --- a system for forming statements, combining them with
connectives, and determining their truth mechanically.
Propositional logic provides that foundation.

\medskip
\noindent\textbf{What this chapter builds.}
We develop the syntax of propositional formulas, their semantics via
truth assignments, and the proof theory of natural deduction.
The central results --- soundness, completeness, and compactness ---
characterise what is provable and what follows logically.

\medskip
\noindent\textbf{Where this leads.}
Predicate logic extends this language with quantifiers, allowing us to
talk about objects and their properties rather than just truth values.
The metatheoretic pattern (syntax, semantics, soundness, completeness)
established here repeats at every level of the logical hierarchy.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Structural Roadmap
% =========================================================
\section*{Structural Roadmap}

Each major topic is organised as:
\begin{center}
\textbf{Syntax $\longrightarrow$ Semantics
$\longrightarrow$ Proof Theory $\longrightarrow$ Metatheory}
\end{center}

The global progression is:
\begin{enumerate}
  \item Syntax: propositional variables, connectives, well-formed formulas
  \item Semantics: truth assignments, validity, logical equivalence, normal forms
  \item Proof theory: inference rules, natural deduction, derivability
  \item Metatheory: functional completeness, compactness, Craig's interpolation
  \item Reference: fallacies and summary tables
\end{enumerate}

\vspace{1em}

\begin{remark}[Primary source]
The primary driver is Bjørndahl's \textit{Logic and Proof},
supplemented by Suppes' \textit{Introduction to Logic}.
\end{remark}

% =========================================================
% Content
% =========================================================
\input{volume-i/propositional-logic/notes/index}
