\subsubsection*{Summary}

\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Toolkit}},
  fonttitle=\small\bfseries
]
This summary collects the normal-form vocabulary: literals, clauses, terms,
NNF, DNF, CNF, canonical forms, and conversion workflow.
\end{tcolorbox}

\begin{remark*}[Exposition]
The normal-forms section turns arbitrary propositional formulas into regular
syntactic objects that are easier to inspect. The guiding invariant throughout
the section is logical equivalence: a conversion may change the shape of a
formula, but it must not change its truth value under any valuation.
\end{remark*}

\begin{remark*}[Literal vocabulary]
The vocabulary begins at the literal level. A literal is a propositional
variable or the negation of one propositional variable. Positive literals and
negative literals record the sign of that occurrence. Complementary literals
are the opposed pair \(P\) and \(\neg P\). Clauses are finite disjunctions of
literals, while terms are finite conjunctions of literals. The empty clause is
the false boundary case, and the empty conjunction is the true boundary case.
\end{remark*}

\begin{remark*}[Negation normal form]
Negation normal form is the first normalization target. It permits only
\(\neg\), \(\land\), and \(\lor\), with every negation pushed down to the
literal level. Its purpose is to remove conditionals and biconditionals and to
make negation local. Once a formula is in NNF, the remaining work is a matter
of distributing \(\land\) and \(\lor\) in the direction required by the target
normal form.
\end{remark*}

\begin{remark*}[Disjunctive normal form]
Disjunctive normal form writes a formula as a finite disjunction of terms. It
is naturally witness-oriented: to make a DNF formula true, it is enough to make
one term true, and a satisfiable term directly records compatible literal
requirements. This is why DNF is especially useful for satisfiability checks
and for building formulas from true rows of a truth table.
\end{remark*}

\begin{remark*}[Conjunctive normal form]
Conjunctive normal form writes a formula as a finite conjunction of clauses. It
is naturally obstruction-oriented: to falsify a CNF formula, one must falsify a
clause, and each clause gives a local disjunctive condition. This clause
structure makes CNF the preferred shape for later clause checking,
resolution-style arguments, and systematic proof search.
\end{remark*}

\begin{remark*}[Canonical normal forms]
Ordinary DNF and CNF are not unique. Reordering, duplicating, or simplifying
terms and clauses can produce different formulas with the same truth table.
Canonical normal forms add truth-table discipline. Canonical DNF records the
true rows by minterm-style conjunctions, and canonical CNF records the false
rows by maxterm-style clauses. These canonical forms show that every finite
truth behavior can be represented syntactically.
\end{remark*}

\begin{remark*}[Conversion workflow]
The conversion workflow has a common beginning. First eliminate
biconditionals. Then eliminate conditionals. Then push negations inward until
the formula is in negation normal form. From there, distribute conjunction over
disjunction to obtain DNF, or distribute disjunction over conjunction to obtain
CNF. Each stage is justified by logical equivalence laws and by replacement of
logical equivalents inside larger formulas.
\end{remark*}

\begin{remark*}[Back to the algebra of propositions]
Normal forms are an application of the algebra of propositions. The conversion
rules are not new semantic principles; they are uses of earlier equivalence
laws such as De Morgan laws, conditional and biconditional equivalences,
distributivity, and replacement of equivalents. The normal-form viewpoint
therefore shows how algebraic manipulation and semantic preservation work
together.
\end{remark*}

\begin{remark*}[Forward to functional completeness]
Canonical normal forms point forward to functional completeness. If a truth
function can be described by its true rows, then it can be represented by a
disjunction of conjunctions of literals. If it can be described by its false
rows, then it can be represented by a conjunction of disjunctions of literals.
This is the bridge from truth tables to the question of which connectives are
enough to express every truth function.
\end{remark*}

\begin{remark*}[Forward to arguments and proof systems]
Normal forms also prepare the transition from single formulas to arguments and
proof systems. DNF gives explicit witnesses for satisfiability, while CNF turns
formulas into collections of clauses that can be checked and transformed. The
same syntactic regularity used here to classify formulas will later support
validity tests, derivation systems, resolution methods, and comparisons between
semantic consequence and formal proof.
\end{remark*}
