% =========================================================
% Structural Induction (Preview)
% =========================================================
\paragraph{Structural Induction (Preview).}

Induction on $\mathbb{N}$ is a special case of a general principle
that applies to any recursively defined structure. You do not need
$\mathbb{N}$ to do induction; you need a recursive definition.

\begin{remark}[The general principle]
Let $X$ be a set defined by:
\begin{enumerate}[itemsep=2pt]
  \item A collection of \emph{base cases} (atomic elements not built
    from other elements), and
  \item A collection of \emph{constructor rules} (ways of building
    larger elements from smaller ones).
\end{enumerate}
Structural induction on $X$: to prove $P(x)$ for all $x \in X$,
prove $P$ for all base cases, then prove that if $P$ holds for the
parts, it holds for the whole.

This is exactly standard induction where the ``less than'' ordering
on $\mathbb{N}$ is replaced by the ``proper sub-element'' ordering
on $X$.
\end{remark}

\begin{example}[Induction on algebraic terms]
Define an \emph{algebraic term} over a ring $R$ by:
\begin{itemize}[itemsep=2pt]
  \item Base case: any element $r \in R$ is a term.
  \item Constructor: if $s, t$ are terms, so are $s + t$ and $s \cdot t$.
\end{itemize}
To prove a property $P$ holds for all terms:
\begin{enumerate}[itemsep=2pt]
  \item Prove $P(r)$ for all $r \in R$ (base cases).
  \item Prove: if $P(s)$ and $P(t)$ hold, then $P(s + t)$ holds.
  \item Prove: if $P(s)$ and $P(t)$ hold, then $P(s \cdot t)$ holds.
\end{enumerate}
The recursive structure of terms determines the structure of the proof.
\end{example}

\begin{remark}[Where structural induction appears in this project]
Induction on $\mathbb{N}$ is structural induction on the Peano
structure $(\mathbb{N}, 0, S)$: the base case is $n = 0$, the
constructor is $n \mapsto n\pp$.

In the Abstract Algebra chapter, polynomial properties are proved by
induction on degree, which is structural induction on the recursive
definition of polynomials over a ring.

In computer science, correctness of recursive algorithms is proved by
structural induction on the input data structure (lists, trees,
expressions). The full treatment of structural induction appears in
the Abstract Algebra chapter.
\end{remark}
