\subsection*{Consistency}
\label{sec:axiom-systems-consistency}

\begin{toolkitbox}{Consistency Toolkit --- Quick Reference}
\small
\begin{tabular}{l l l}
\toprule
\textbf{Concept} & \textbf{Role} & \textbf{Detail} \\
\midrule
Contradiction & assertion that cannot be true in an interpretation & \hyperref[def:axiom-systems-contradiction]{Definition} \\
Consistency & absence of contradiction & \hyperref[def:axiom-systems-consistent-axiom-system]{Definition} \\
Inconsistency & contradiction in or from axioms & \hyperref[def:axiom-systems-inconsistent-axiom-system]{Definition} \\
Model existence & evidence that axioms fit together & \hyperref[def:axiom-systems-model-existence-criterion]{Definition} \\
Relative consistency & consistency transferred from another theory & \hyperref[def:axiom-systems-relative-consistency]{Definition} \\
\bottomrule
\end{tabular}
\end{toolkitbox}

\begin{tcolorbox}[colback=gray!6, colframe=gray!40, arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Consistency Presentation: A Simple Example}},
  fonttitle=\small\bfseries]
Consider two informal axiom systems.

\medskip
\noindent\textbf{Example A.}
\begin{itemize}
  \item All objects are red.
  \item Some object is not red.
\end{itemize}
These assertions cannot simultaneously hold. The first assertion says every
object has the property; the second says at least one object lacks it.

\medskip
\noindent\textbf{Example B.}
\begin{itemize}
  \item All objects are red.
  \item Object $a$ is red.
\end{itemize}
These assertions can hold simultaneously. The second assertion agrees with
what the first assertion already requires.

\medskip
These examples are only for intuition. A consistent axiom system is one whose
assertions fit together without contradiction. An inconsistent axiom system is
one whose assertions cannot be held together coherently.
\end{tcolorbox}

\begin{remark*}[Structural remarks]
\textbf{Why contradictions matter.}
Contradictory systems are unusable because their assertions do not fit
together. If a system both requires and forbids the same situation, then it no
longer gives a coherent basis for reasoning.

\textbf{Models as evidence.}
If a model exists, the axioms hold together coherently in at least one
interpretation. A model is therefore evidence that the axioms are mutually
compatible.

\textbf{Absolute vs relative consistency.}
Most consistency results in mathematics are relative consistency results. They
show that one theory is consistent provided some other background theory is
consistent. This section records the idea only; the technical development
belongs elsewhere.

\textbf{Forward roadmap.}
The next section uses models and consistency to introduce independence:
\[
\begin{array}{c}
\text{Consistency}\\
\Downarrow\\
\text{Independence}
\end{array}
\]
\end{remark*}

\begin{definition}[Contradiction]
\label{def:axiom-systems-contradiction}
A \emph{contradiction} is a statement that cannot be true under a given
interpretation.
\end{definition}

\begin{remark*}[Interpretation]
A contradiction marks a breakdown in compatibility. It identifies an assertion
that cannot be made true in the interpretation under consideration.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:axiom-systems-satisfaction]{Satisfaction}
\end{itemize}
\end{dependencies}

\begin{definition}[Consistent Axiom System]
\label{def:axiom-systems-consistent-axiom-system}
A \emph{consistent axiom system} is an axiom system that does not lead to
contradiction.
\end{definition}

\begin{remark*}[Interpretation]
Consistency says that the adopted assertions fit together. A consistent axiom
system avoids forcing an impossible situation.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:axiom-system]{Axiom System}
  \item \hyperref[def:axiom-systems-contradiction]{Contradiction}
\end{itemize}
\end{dependencies}

\begin{definition}[Inconsistent Axiom System]
\label{def:axiom-systems-inconsistent-axiom-system}
An \emph{inconsistent axiom system} is an axiom system containing or producing
contradictory assertions.
\end{definition}

\begin{remark*}[Interpretation]
Inconsistency means that the adopted assertions cannot all be maintained
coherently. The system asks for incompatible requirements at once.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:axiom-system]{Axiom System}
  \item \hyperref[def:axiom-systems-contradiction]{Contradiction}
\end{itemize}
\end{dependencies}

\begin{definition}[Model Existence Criterion (Informal)]
\label{def:axiom-systems-model-existence-criterion}
Informally, if an axiom system has a model, then the axioms are mutually
compatible.
\end{definition}

\begin{remark*}[Interpretation]
This is an informal criterion in this section, not a formal theorem. The idea
is that a model provides a setting in which all the axioms hold together, so
the axioms have at least one coherent realization.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:axiom-systems-model]{Model}
\end{itemize}
\end{dependencies}

\begin{definition}[Relative Consistency]
\label{def:axiom-systems-relative-consistency}
A \emph{relative consistency} statement says that one theory is consistent
provided another theory is consistent.
\end{definition}

\begin{remark*}[Interpretation]
Relative consistency does not establish consistency from nothing. It compares
systems: if the background theory avoids contradiction, then the target theory
does as well.
\end{remark*}

\begin{dependencies}
\begin{itemize}
  \item \hyperref[def:axiom-systems-consistent-axiom-system]{Consistent Axiom System}
\end{itemize}
\end{dependencies}
