% =========================================================
% Formula Depth (Complexity)
% =========================================================

\section{Syntax: Formula Depth}

% ---------------------------------------------------------
% TOOLKIT
% ---------------------------------------------------------
\begin{tcolorbox}[colback=gray!6, colframe=gray!40, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Formula Depth — Quick Reference}},
  fonttitle=\small\bfseries]
\small
\begin{tabular}{l l l}
\toprule
\textbf{Concept} & \textbf{Meaning} & \textbf{Detail} \\
\midrule
Formula depth & Max formation steps from atomic formulas & \hyperref[def:formula-depth]{↓ Def} \\
\bottomrule
\end{tabular}
\end{tcolorbox}

\vspace{1em}

% ---------------------------------------------------------
% Formula Depth
% ---------------------------------------------------------
\begin{tcolorbox}[colback=propbox, colframe=propborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Definition (Formula Depth)}},
  fonttitle=\small\bfseries]
\label{def:formula-depth}
The \emph{depth} (or \emph{complexity}) of a formula $\varphi$, denoted
$\mathrm{depth}(\varphi)$, is a natural number defined recursively:
\begin{enumerate}
  \item \textbf{Atomic case.} If $\varphi$ is atomic, then $\mathrm{depth}(\varphi) = 0$.

  \item \textbf{Negation.} If $\varphi = \neg \psi$, then
  $\mathrm{depth}(\varphi) = \mathrm{depth}(\psi) + 1$.

  \item \textbf{Binary connectives.} If $\varphi = (\psi \circ \chi)$
  for $\circ \in \{\wedge,\vee,\rightarrow,\leftrightarrow\}$, then
  \[
  \mathrm{depth}(\varphi) = \max\{\mathrm{depth}(\psi),\,\mathrm{depth}(\chi)\} + 1.
  \]

  \item \textbf{Quantifiers.} If $\varphi = \forall x\,\psi$ or $\varphi = \exists x\,\psi$,
  then $\mathrm{depth}(\varphi) = \mathrm{depth}(\psi) + 1$.
\end{enumerate}
\end{tcolorbox}

\begin{remark}[English reading]
The depth of a formula measures the maximum number of logical formation steps
required to build the formula from atomic formulas. It corresponds to the height
of the formula's syntactic parse tree.
\end{remark}

\begin{remark}[Proof strategy]
Formula depth is the standard measure used in structural induction proofs over
first-order formulas. The base case is depth~$0$ (atomic formulas); the
inductive step handles each formation rule and assumes the result holds for
sub-formulas of strictly smaller depth.
\end{remark}
