% =========================================================
% Proof by Cases
% =========================================================
\paragraph{Proof by Cases.}

If the hypotheses or conclusion naturally divide into a finite number
of mutually exclusive and collectively exhaustive alternatives, handle
each case separately.

\begin{tcolorbox}[colback=propbox, colframe=propborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Template: Proof by Cases}}, fonttitle=\small\bfseries]
\textbf{Goal:} Prove $Q$, given that exactly one of $P_1, P_2, \ldots, P_k$
holds (and one of them always holds).

\medskip
\textit{Verify exhaustiveness:} Every possibility falls under some $P_i$.\\
\textit{Verify mutual exclusivity} (when needed).

\medskip
\textbf{Case 1:} Assume $P_1$. \quad [Argue.] \quad Hence $Q$.\\[4pt]
\textbf{Case 2:} Assume $P_2$. \quad [Argue.] \quad Hence $Q$.\\[4pt]
\hspace{1.2em}$\vdots$\\[4pt]
\textbf{Case $k$:} Assume $P_k$. \quad [Argue.] \quad Hence $Q$.

\medskip
Since $P_1, \ldots, P_k$ are exhaustive and $Q$ holds in each case,
$Q$ holds unconditionally. \qed
\end{tcolorbox}

\begin{remark}[The exhaustiveness obligation is non-negotiable]
A case proof is only valid when the cases together cover every
possibility. Always verify exhaustiveness explicitly, even when it
feels obvious.

The most common exhaustive splits:
\begin{itemize}[itemsep=2pt]
  \item $n$ is even, or $n$ is odd.
  \item $a > 0$, $a = 0$, or $a < 0$.
  \item $a \mid b$, or $a \nmid b$.
  \item $x \in A$, or $x \notin A$.
  \item $x = y$, or $x \neq y$.
\end{itemize}

Writing ``we consider two cases'' without naming what they are, or
naming cases that do not cover all possibilities, is a structural error.
The proof is not valid until exhaustiveness is established.
\end{remark}

\begin{remark}[Each case is a self-contained argument]
Inside each case, you have an additional assumption: $P_i$ is true.
That assumption is local to the case. You may use it within Case~$i$,
but you may not carry it into Case~$j$.

When you close a case --- ``hence $Q$, completing Case~1'' --- you
are signalling that the local assumption is no longer in scope.
Case~2 starts fresh, with only the original given plus the assumption $P_2$.

This discipline prevents the most common case-analysis error: drifting
from one case into another by accidentally retaining a local assumption
beyond its scope.
\end{remark}

\begin{remark}[Symmetric cases]
When two cases are exactly symmetric --- when the argument for Case~2
is obtained from the argument for Case~1 by renaming variables, with
no other change --- it is acceptable to write ``Case~2 is symmetric
to Case~1 by swapping $a$ and $b$.'' But be careful: only write this
when the symmetry is complete. A partial symmetry that requires a
different step at one point must be written out in full.
\end{remark}

\begin{example}[Parity of $n(n+1)$]
\textbf{Claim.} For every $n \in \mathbb{Z}$, $n(n+1)$ is even.

\medskip
\noindent\textit{Proof.}
Every integer is either even or odd (these cases are exhaustive and
mutually exclusive).

\textbf{Case 1: $n$ is even.}
Write $n = 2k$ for some $k \in \mathbb{Z}$.
Then $n(n+1) = 2k(n+1)$. Since this is $2$ times an integer,
$n(n+1)$ is even.

\textbf{Case 2: $n$ is odd.}
Write $n = 2k+1$ for some $k \in \mathbb{Z}$.
Then $n+1 = 2k+2 = 2(k+1)$, so $n(n+1) = (2k+1) \cdot 2(k+1)$.
This is $2$ times an integer, so $n(n+1)$ is even.

In both cases $n(n+1)$ is even. Since every integer falls into one of
these cases, the proof is complete. \qed

\medskip
\noindent
Notice that the cases were named before the argument, the exhaustiveness
was verified (``every integer is even or odd''), and each case was
closed explicitly before moving to the next.
\end{example}

\begin{example}[Case split induced by a definition]
\textbf{Claim.} The discrete metric satisfies the triangle inequality.

\medskip
\noindent\textit{Proof.}
Let $x, y, z \in X$. We must show $d(x,z) \leq d(x,y) + d(y,z)$.

\textbf{Case 1: $x = z$.}
Then $d(x,z) = 0$. Since $d(x,y) \geq 0$ and $d(y,z) \geq 0$,
we have $0 \leq d(x,y) + d(y,z)$.

\textbf{Case 2: $x \neq z$.}
Then $d(x,z) = 1$. We cannot have both $x = y$ and $y = z$,
for that would force $x = z$. So at least one of $d(x,y), d(y,z)$
equals~$1$, hence $d(x,y) + d(y,z) \geq 1 = d(x,z)$.

The cases $x = z$ and $x \neq z$ are exhaustive. The triangle
inequality holds in both cases. \qed

\medskip
\noindent
The case split here is induced by the two-value structure of the
discrete metric: distances are $0$ or $1$. The right case split
to use is the one naturally suggested by the definition you are working with.
\end{example}
