% =========================================================
% Proof Strategies: From Goal to Method
% =========================================================

\section{Proof Strategies: Connecting Logic to Mathematical Practice}

% ---------------------------------------------------------
% TOOLKIT
% ---------------------------------------------------------
\begin{tcolorbox}[colback=gray!6, colframe=gray!40, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Proof Strategy Guide — Quick Reference}},
  fonttitle=\small\bfseries]
\small
\begin{tabular}{l l l}
\toprule
\textbf{Goal shape} & \textbf{Primary strategy} & \textbf{Key rule} \\
\midrule
Prove $P \to Q$              & Assume $P$, derive $Q$, discharge          & Conditional Intro (CP) \\
Prove $P \to Q$ (indirectly) & Assume $\neg Q$, derive $\neg P$           & Contraposition \\
Prove $\neg P$               & Assume $P$, derive $\bot$, discharge       & Negation Intro (RAA) \\
Prove $P$ (from $\neg P \to \bot$) & Assume $\neg P$, derive $\bot$         & Indirect Proof \\
Prove $P \wedge Q$           & Prove $P$ and $Q$ separately               & Conjunction Intro \\
Prove $P \vee Q$             & Prove $P$, or prove $Q$, or use a split    & Disjunction Intro \\
Prove $P \leftrightarrow Q$  & Prove $P \to Q$ and $Q \to P$ separately   & Biconditional Intro \\
Use $P \vee Q$               & Consider case $P$ and case $Q$ separately  & Disjunction Elim \\
Use $P \to Q$ and $P$        & Apply directly                             & Modus Ponens \\
Use $P \to Q$ and $\neg Q$   & Derive $\neg P$                            & Modus Tollens \\
\bottomrule
\end{tabular}
\end{tcolorbox}

\vspace{1em}

The inference rules of the previous section are the atoms of formal reasoning:
each rule licenses a single step. But in practice, one never asks ``which
single rule applies here?'' One asks: ``what is the shape of my goal, and what
strategy does that shape call for?'' The goal dictates the strategy; the
strategy selects the rules. This section makes that mapping explicit.

% ---------------------------------------------------------
% The anatomy of a proof
% ---------------------------------------------------------

\section*{The Anatomy of a Proof}

Every formal proof has the same three-part structure. Understanding this
structure before attempting any proof prevents the most common errors.

\begin{tcolorbox}[colback=propbox, colframe=propborder, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{The Three-Part Structure of a Proof}},
  fonttitle=\small\bfseries]
\begin{enumerate}[label=(\roman*)]
\item \textbf{Available:} The premises and assumptions currently in force.
  These are what you are allowed to use. At the start, this is the hypothesis
  set $\Gamma$. Inside a subproof, additional assumptions are temporarily added.
\item \textbf{Goal:} The formula you are trying to derive. Every proof step
  either moves toward the goal directly or adds something to ``Available'' that
  will be useful later.
\item \textbf{Strategy:} The top-level decision about how to attack the goal.
  This depends almost entirely on the \emph{shape} of the goal formula.
\end{enumerate}
\end{tcolorbox}

The most important principle is: \emph{look at the goal first, not the
hypotheses}. Beginners often scan their hypotheses and start applying rules
without a target in mind. This produces proof fragments that go nowhere. The
goal tells you what kind of proof you need; the hypotheses tell you what
material you have to work with.

% ---------------------------------------------------------
% Proving a conditional
% ---------------------------------------------------------

\section*{Strategy 1: Proving a Conditional $P \to Q$}

If the goal is $P \to Q$, the standard approach is \emph{conditional proof}
(also called \emph{direct proof} of a conditional): assume $P$ temporarily,
derive $Q$ under that assumption, then discharge the assumption and conclude
$P \to Q$ by Conditional Introduction. The proof skeleton is:

\begin{quote}
\textbf{Assume:} $P$. \\
\textbf{[derive $Q$ from the available hypotheses and $P$]} \\
\textbf{Conclude:} $P \to Q$. \quad (by Conditional Introduction)
\end{quote}

The assumption $P$ is in force only within the subproof; it is discharged at
the moment Conditional Introduction is applied. Nothing derived inside the
subproof using $P$ may be exported outside it, except $P \to Q$ itself.

\begin{example}[Direct conditional proof]
Prove: $(P \wedge Q) \to P$.

\begin{enumerate}
\item Assume $P \wedge Q$.
\item From (1), $P$ \quad (by Conjunction Elimination).
\item Therefore $(P \wedge Q) \to P$ \quad (by Conditional Introduction, discharging assumption 1).
\end{enumerate}
\end{example}

% ---------------------------------------------------------
% Contrapositive
% ---------------------------------------------------------

\section*{Strategy 2: Proof by Contrapositive}

If the goal is $P \to Q$ and the direct route (assuming $P$, deriving $Q$) is
not obvious, consider instead proving $\neg Q \to \neg P$, which is logically
equivalent by contraposition ($P \to Q \equiv \neg Q \to \neg P$).

\begin{quote}
\textbf{Assume:} $\neg Q$. \\
\textbf{[derive $\neg P$ from the available hypotheses and $\neg Q$]} \\
\textbf{Conclude:} $\neg Q \to \neg P$, hence $P \to Q$. \quad (by Contraposition)
\end{quote}

\begin{remark}[When contrapositive is preferred]
Proof by contrapositive is typically chosen when the hypothesis $\neg Q$ is
richer or easier to work with than $P$. A common situation in analysis: to
prove ``if $f$ is continuous at $a$, then $f$ satisfies property $Q$,'' it is
sometimes easier to assume $f$ fails property $Q$ (i.e., $\neg Q$) and derive
that $f$ is not continuous at $a$ (i.e., $\neg P$).
\end{remark}

% ---------------------------------------------------------
% Contradiction
% ---------------------------------------------------------

\section*{Strategy 3: Proof by Contradiction}

There are two related but distinct uses of contradiction. They have different
goals.

\medskip
\noindent\textbf{Proving $\neg P$ (Negation Introduction / RAA):}

If the goal is $\neg P$, assume $P$ temporarily, derive a contradiction $\bot$
(typically by deriving both some $Q$ and $\neg Q$), and conclude $\neg P$ by
Negation Introduction.

\begin{quote}
\textbf{Assume:} $P$. \\
\textbf{[derive $Q$]} \\
\textbf{[derive $\neg Q$ from the same assumptions]} \\
\textbf{Conclude:} $\bot$, then $\neg P$. \quad (by Negation Introduction)
\end{quote}

\medskip
\noindent\textbf{Proving $P$ (Indirect Proof / Negation Elimination):}

If the goal is $P$ itself and a direct route is not apparent, assume $\neg P$
temporarily, derive a contradiction, and conclude $P$ by Negation Elimination
(Indirect Proof). This is the classical reductio: one shows $\neg P$ is
untenable.

\begin{quote}
\textbf{Assume:} $\neg P$. \\
\textbf{[derive a contradiction $\bot$]} \\
\textbf{Conclude:} $P$. \quad (by Negation Elimination)
\end{quote}

\begin{remark}[When to use contradiction vs. contrapositive]
Both strategies involve a negated assumption, so they are easy to confuse.
The difference is the goal: contradiction is used when the goal is hard to
reach directly and it is easier to show its negation leads to absurdity.
Contrapositive is used specifically for a conditional goal $P \to Q$.
In practice, when the goal is $\neg P$, use Negation Introduction (RAA). When
the goal is an arbitrary $P$ and you are stuck, try Indirect Proof (assume
$\neg P$).
\end{remark}

\begin{remark}[Classical logic caveat]
Indirect Proof (Negation Elimination) and the equivalence $P \equiv \neg\neg P$
are classical logic rules. They fail in intuitionistic logic, where proving $\neg P$
is impossible leads to absurdity but does not, on its own, establish $P$.
All proofs in this text use classical logic.
\end{remark}

% ---------------------------------------------------------
% Proof by cases
% ---------------------------------------------------------

\section*{Strategy 4: Proof by Cases (Disjunction Elimination)}

If a hypothesis or known fact has the form $P \vee Q$, one cannot simply use
$P$ or use $Q$ --- neither is known on its own. Instead, split into two
subproofs: prove the goal $R$ in the case where $P$ holds, and prove $R$ in
the case where $Q$ holds. Since one of $P$ or $Q$ must be true, $R$ follows
in either case.

\begin{quote}
\textbf{Available:} $P \vee Q$. Goal: $R$. \\
\textbf{Case 1:} Assume $P$. \quad [derive $R$] \\
\textbf{Case 2:} Assume $Q$. \quad [derive $R$] \\
\textbf{Conclude:} $R$. \quad (by Disjunction Elimination)
\end{quote}

\begin{remark}[Cases need not be $P$ and $\neg P$]
Proof by cases is often written as ``case $A$ holds'' and ``case $A$ does not
hold'' — an exhaustive split. This is a special case where $P \vee Q$ is
$A \vee \neg A$ (the law of excluded middle). But cases can come from any
disjunction in the hypotheses, not just LEM.
\end{remark}

\begin{example}[Case split on an available disjunction]
Suppose the hypotheses include $P \vee Q$, $P \to R$, and $Q \to R$.
Goal: $R$.

\begin{enumerate}
\item From $P \to R$, $Q \to R$, and $P \vee Q$: apply Constructive Dilemma
  to obtain $R \vee R$.
\item From $R \vee R$: apply idempotence or Disjunction Elimination to obtain $R$.
\end{enumerate}

Alternatively, use Disjunction Elimination directly: in case $P$, apply MP to
get $R$; in case $Q$, apply MP to get $R$; conclude $R$.
\end{example}

% ---------------------------------------------------------
% Biconditional strategy
% ---------------------------------------------------------

\section*{Strategy 5: Proving a Biconditional $P \leftrightarrow Q$}

If the goal is $P \leftrightarrow Q$, always split into two sub-goals using
Biconditional Introduction:

\begin{quote}
\textbf{Part ($\Rightarrow$):} Prove $P \to Q$ (using Conditional Introduction).\\
\textbf{Part ($\Leftarrow$):} Prove $Q \to P$ (using Conditional Introduction).\\
\textbf{Conclude:} $P \leftrightarrow Q$. \quad (by Biconditional Introduction)
\end{quote}

This is the standard structure of every ``if and only if'' proof in mathematics.
The two directions are typically argued separately and may require different
strategies.

% ---------------------------------------------------------
% Decision flowchart
% ---------------------------------------------------------

\section*{Putting It Together: The Strategy Decision Process}

The following sequence of questions, applied to the current proof goal, leads
to the right strategy in nearly every case.

\begin{tcolorbox}[colback=gray!6, colframe=gray!40, arc=2pt,
  left=6pt, right=6pt, top=4pt, bottom=4pt,
  title={\small\textbf{Strategy Selection: Questions to Ask}},
  fonttitle=\small\bfseries]
\small
\begin{enumerate}[label=\textbf{Q\arabic*.}]
\item \textbf{Is the goal a conditional $P \to Q$?}
  If yes: assume $P$ (conditional proof). If that seems hard, try assuming
  $\neg Q$ (contrapositive) instead.

\item \textbf{Is the goal a negation $\neg P$?}
  If yes: assume $P$, derive a contradiction (Negation Introduction / RAA).

\item \textbf{Is the goal a biconditional $P \leftrightarrow Q$?}
  If yes: split into two conditional proofs.

\item \textbf{Is the goal a conjunction $P \wedge Q$?}
  If yes: prove $P$ and prove $Q$ independently, then apply Conjunction Introduction.

\item \textbf{Is a disjunction $P \vee Q$ available as a hypothesis?}
  If yes: consider proof by cases (Disjunction Elimination).

\item \textbf{Are you stuck with a non-connective goal?}
  If yes: try Indirect Proof (assume the negation of the goal, derive a
  contradiction).
\end{enumerate}
\end{tcolorbox}

\begin{remark}[The order matters]
Apply these questions top-to-bottom. The goal shape (Q1--Q4) is checked before
the hypothesis shape (Q5). This order reflects the fundamental principle: the
goal drives the proof structure; hypotheses supply the material.
\end{remark}

\begin{remark}[Connection to mathematical practice]
Every standard proof technique in real analysis is an instance of one of these
five strategies:
\begin{itemize}
\item Most analysis proofs prove a conditional (``if $\varepsilon > 0$ then
  $\exists \delta > 0$ such that \ldots'') by conditional proof: fix
  $\varepsilon > 0$ and construct $\delta$.
\item Proofs of irrationality (``$\sqrt{2}$ is irrational'') use Indirect
  Proof: assume $\sqrt{2}$ is rational, derive a contradiction.
\item Proofs about sets often use case splits: either $x \in A$ or $x \notin A$.
\item Proofs of ``iff'' characterizations use biconditional introduction: two
  separate directional arguments.
\end{itemize}
The logical strategies learned here are the skeleton of every proof that will
be written from Volume II onward.
\end{remark}
