% =========================================================
% Breadcrumb
% =========================================================
\begin{tcolorbox}[
  colback=gray!6,
  colframe=gray!40,
  arc=2pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  title={\small\textbf{Where You Are in the Journey}},
  fonttitle=\small\bfseries
]
\begin{center}
\small
Propositional Logic
$\;\to\;$ Predicate Calculus
$\;\to\;$ Sets \& Functions
$\;\to\;$ Proof Techniques
$\;\to\;$ \textbf{Axiom Systems}
$\;\to\;$ Natural Numbers ($\mathbb{N}$)
$\;\to\;$ $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$
$\;\to\;$ $\cdots$
\end{center}

\medskip
\noindent\textbf{How we got here.}
Proof techniques gave us the tools to reason rigorously --- direct proof,
contradiction, induction.
We now turn those tools on mathematics itself: can we build the natural
numbers from nothing but logical axioms?
This chapter answers yes.

\medskip
\noindent\textbf{What this chapter builds.}
We lay down the five Peano axioms, which pin down the structure
of zero and the successor function.
We verify that small numerals ($1, 2, 3, \ldots$) exist, confirm
that the axioms rule out pathological models, and prove the
Recursion Theorem --- the logical guarantee that recursive
definitions actually produce well-defined functions.
These are the \emph{logical primitives}: the minimum needed for
$\mathbb{N}$ to exist.

\medskip
\noindent\textbf{Where this leads.}
Volume~II, Natural Numbers takes over immediately: it uses the Peano
axioms and the Recursion Theorem as its foundation and builds the full
arithmetic of $\mathbb{N}$ --- addition, order, multiplication, and
exponentiation --- from scratch.
Nothing from that development is assumed here.
\end{tcolorbox}
\vspace{1em}

% =========================================================
% Structural Roadmap
% =========================================================
\section*{Structural Roadmap}

This chapter covers the \emph{logical skeleton} of $\mathbb{N}$ only.
Arithmetic (addition, order, multiplication, exponentiation) lives in
\textbf{Volume~II, Natural Numbers}.

\begin{center}
\textbf{Axioms $\longrightarrow$ Numerals
$\longrightarrow$ Induction principle $\longrightarrow$ Recursion theorem}
\end{center}

The global progression is:
\begin{enumerate}
  \item \textbf{Peano axioms} (P1--P5): the five axioms that characterise
        $\mathbb{N}$ up to isomorphism. Includes the induction axiom
        and a discussion of what the axioms exclude.
  \item \textbf{Numerals}: defining $1 := 0\pp$, $2 := 1\pp$, \ldots{}
        and verifying they are distinct natural numbers.
  \item \textbf{Recursion theorem} (Tao Prop.~2.1.16):
        for any set $X$, base value $a \in X$, and function $f : X \to X$,
        there exists a unique $g : \mathbb{N} \to X$ satisfying
        $g(0) = a$ and $g(n\pp) = f(g(n))$.
        This theorem is what makes recursive \emph{definitions} legitimate.
\end{enumerate}

\vspace{1em}

\begin{remark}[Scope boundary]
The Peano axioms tell us $\mathbb{N}$ \emph{exists} and has the right
logical shape. The Recursion Theorem tells us recursive definitions
are \emph{valid}. Everything else --- what $\mathbb{N}$ can \emph{do}
arithmetically --- is in Volume~II.
\end{remark}

\begin{remark}[Primary source]
Tao, \textit{Analysis~I}, Chapter~2, §2.1.
\end{remark}

% =========================================================
% Content
% =========================================================
\input{volume-i/axiom-systems/notes/peano/index}
\input{volume-i/axiom-systems/notes/von-neuman/index}
\input{volume-i/axiom-systems/notes/zermelo/index}

\input{volume-i/axiom-systems/notes/presburger/index}

